---
title: "Workshop - January 2018: Birthday Problem"
author: "cmachingauta"
date: "January 10, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
 # fixing seed 
n.pers =27
set.seed(1234)
```

# The Problem
What is the probability that at least two people this room share the same birthday?

Is it something like $\frac {`r n.pers`} {365} =$ 
`r round(n.pers/365, 2)`

# Analytical solution

\begin{align}
1 - \bar p(n) &= 1 \times \left(1-\frac{1}{365}
\right)
\times \left(1-\frac{2}{365}\right) \times \cdots
\times
\left(1-\frac{n-1}{365}\right) \nonumber \\
&= \frac{ 365 \times 364 \times \cdots \times
(365-n+1) }{ 365^n } \nonumber \\
&= \frac{ 365! }{ 365^n (365-n)!} =
\frac{n!\cdot\binom{365}{n}}{365^n}\\
p(n= `r n.pers`) &= `r round(1 - factorial(n.pers)*choose(365,n.pers)/ 365^n.pers, 3)`
\nonumber
\end{align}

# Simulation

1 - Simulate 10,000 rooms with n = `r n.pers` random birthdays, and store the results in matrix where each row represents a room.   

2 - For each room (row) compute the number of unique birthdays.   

3 - Compute the average number of times a room has `r n.pers` unique birthdays, across 10,000 simulations, and report the complement.   


```{r simulation}
birthday.prob = function(n.pers, n.sims) {  
# simulate birthdays  
birthdays = matrix(round(runif(n.pers * n.sims,1
                               , 365)),nrow = n.sims, 
                   ncol = n.pers)
# for each room (row) get unique birthdays
unique.birthdays = apply(birthdays, 1, unique)  
# Indicator with 1 if all are unique birthdays
all.different = (lapply(unique.birthdays,length) == n.pers)  
# Compute average time all have different birthdays
result = 1 - mean(all.different)  
return(result)  
}
n.pers.param = n.pers  
n.sims.param = 1e4  
res1 = birthday.prob(n.pers.param,n.sims.param)
```

The simulated probability is `r round(res1, 2)`

